\subsection{Considerations}

\begin{frame}
    \frametitle{Considerations}

    This work presents the development and analysis of the \textit{k-sGMRES} algorithm, an efficient variant of the \textit{GMRES} method for solving large linear systems. Key observations include:

    \begin{itemize}
        \item The algorithm demonstrates superior performance compared to \lstinline{MATLAB}'s \textit{GMRES}:
        \begin{itemize}
            \item Comparable accuracy in terms of relative residuals.
            \item Computational time reduced to nearly one-hundredth of \lstinline{MATLAB}'s runtime.
        \end{itemize}
        \item \textit{k-sGMRES} can handle arbitrarily large matrices, with the only constraint being the available memory or storage.
    \end{itemize}

    These characteristics make \textit{k-sGMRES} highly suitable for large-scale problems requiring efficiency and accuracy.
\end{frame}

\subsection{Possible Improvements}

\begin{frame}
    \frametitle{Possible Improvements}

    Future enhancements to the \textit{k-sGMRES} implementation include:

    \begin{itemize}
        \item \textbf{Support for Larger Matrices}
        \begin{itemize}
            \item Utilize direct storage access to reduce memory overhead, shifting the constraint from memory to storage capacity.
            \item Employ matrix handlers to enable streamed matrix constructions, optimizing memory usage during computations.
        \end{itemize}

        \item \textbf{Improved Convergence for Ill-Conditioned Matrices}
        \begin{itemize}
            \item Apply simple and general preconditioning techniques to enhance convergence, particularly for matrices like \textbf{stokes}.
            \item Advanced, problem-specific preconditioning methods lie outside the scope of this work but represent a promising direction for future research.
        \end{itemize}
    \end{itemize}

\end{frame}